### 2022-02-03 outputs/paper/paper.Rmd
Added introduction

---
title: "Toronto Apartment Evaluations"
subtitle: "Subtitle Here"
author: 
  - Ritvik Puri
  # - Another author
thanks: "Code and data are available at: www.github.com/ritvikpuri/sta304-paper-1"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
# library(palmerpenguins)
library(opendatatoronto)
```

# Introduction

<!-- You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).  -->

The city of Toronto introcued a new bylaw enformenent program in July 2017 that is used to ensure that the apartment building owners and operators comply with the neccesary building maintanence standards, known as RentSafeTO. These standards are applied to all apartment buildings with 10 or more units  or with 3 or more stories.

Owners of such apartments are required to register with RentSafeTO as well as mainitan the standards defined by this program. Tentants need to conatct their respective landlord incase they face an issue. These issues could either be vital, such as heat or hydro fault, or service requests, such as window flaws or common area cleaning.  If the landlord does not comply to these requests, then according to the bylaws of RentSafeTO program, legal action can be taken against them.

Each property that falls under the program gets insepected by an officer and receives an evaluation score, which is made available to not only the landlord and tenant, but also to the potential tenants. If the score of a building is 86 or above, it will be evaluated again in the next three years. If it is between 66 and 85, it will be evaluated again within two years. If it is between 51 and 65, it will be evaluated again within a year. If a building gets a score of 50 or below, then then the full building will undergo a comprehensive inspection.

Home safety evaluations are neccesary to protect residents from potential hazards which may lead to personal injury if left unchecked. In a city like Toronto, where a lot of people do not own their personal property and live in rented apartments, it is critical that the landlord is kept informed of their building's condition so that they can ensure a tenants safety.

# Data

<!-- Our data is of penguins (Figure \@ref(fig:bills)). -->

# ```{r bills, fig.cap="Bills of penguins", echo = FALSE}
# ggplot(penguins, aes(x = island, fill = species)) +
#   geom_bar(alpha = 0.8) +
#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                     guide = FALSE) +
#   theme_minimal() +
#   facet_wrap(~species, ncol = 1) +
#   coord_flip()
# ```

```{r}

package <- show_package("4ef82789-e038-44ef-a478-a8f3590c3eb1")
package

# get all resources for this package
resources <- list_package_resources("4ef82789-e038-44ef-a478-a8f3590c3eb1")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()

view(data)
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
# ggplot(penguins, aes(x = island, fill = species)) +
#   geom_bar(alpha = 0.8) +
#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                     guide = FALSE) +
#   theme_minimal() +
#   facet_wrap(~species, ncol = 1) +
#   coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2022-02-04 outputs/paper/paper.Rmd
First model added

@@
 In a city like Toronto, where a lot of people do not own their personal property and live in rented apartments, it is critical that the landlord is kept informed of their building's condition so that they can ensure a tenants safety.
+Dhekte hein ki kuch badiya sa milta hai ya nahi inn socres ke baare mein.
+Aayie shuru karte, bina kisi rukawat ke.
+
+I think we should see which factors have the most effect on the evaluation score.
 
@@
 <!-- Our data is of penguins (Figure \@ref(fig:bills)). -->
-
-# ```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = FALSE) +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
-# ```
 
@@
 
+sapply(data, class)
+data[,1:5] <- sapply(data[,1:5], as.numeric)
+data[,7:7] <- sapply(data[,7:7], as.numeric)
+data[,10:11] <- sapply(data[,10:11], as.numeric)
+data[,13:13] <- sapply(data[,13:13], as.numeric)
+data[,14:35] <- sapply(data[,14:35], as.numeric)
+
 view(data)
@@
 
-# This needs to be about the random data tha tI sasmpled?
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = FALSE) +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
+#unique proerty types
+unique_properties <- unique(data$PROPERTY_TYPE)
+
+data %>%
+  ggplot(mapping = aes(x = SCORE, y = CONFIRMED_UNITS, color= WARDNAME)) + 
+  geom_point(alpha = 0.5)
+
+data %>%
+  ggplot(mapping = aes(y = SCORE, fill=WARDNAME)) +
+  geom_bar()
+
+mut_data <- data %>%
+  mutate(score_category =
+           case_when(SCORE <= 50 ~ "<=50",
+                     SCORE <= 65 ~ "51-65",
+                     SCORE <= 85 ~ "66-85",
+                     SCORE <= 100 ~ "86-100"),
+         score_category = factor(score_category,
+                                 levels = c("<=50",
+                                            "51-65",
+                                            "66-85",
+                                            "86-100"
+                                            )
+                                 )
+         )
+view(mut_data)
+mut_data %>%
+  ggplot(mapping = aes(x = score_category, fill = PROPERTY_TYPE)) +
+  geom_bar(position = "dodge")
 ```
 
-Talk way more about it.
+We can see how there are a lot of apartment building with a score thats between 66 and 85, which means they wont be evaluated for the next 2 years and are consiered to be in good condition overall.
+But what makes the score of these buildings higher than the rest?
+And what do the few buildings with a score of more than 86 do that separtes them from the rest?
 
@@
 # References
+https://www.toronto.ca/community-people/housing-shelter/rental-housing-tenant-information/rental-housing-standards/apartment-building-standards/rentsafeto-for-tenants/
+
+https://www.toronto.ca/community-people/housing-shelter/rental-housing-tenant-information/rental-housing-standards/apartment-building-standards/rentsafeto-for-building-owners/rentsafeto-building-evaluations-and-audits/

### 2022-02-06 outputs/paper/paper.Rmd
write up almost complete

@@
 ---
-title: "Toronto Apartment Evaluations"
-subtitle: "Subtitle Here"
+title: "Toronto Apartment Evaluations Are Hard to Predict"
+#subtitle: "Subtitle Here"
 author: 
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Building evaluations are always important to conduct, not only from a safety perspective, but also to promote people buying and renting more property.
+The RentSafeTO program started in 2017, and the data is updated daily on Open Data Toronto, and we will be looking at the various factors that play in the role of giving a better score.
+Over the last 5 years, we have seen a generally upward trend in building scores.
+We also find that there are certain features, like multiple stairwells and higher security, that tend to procure a extra points."
 output:
@@
 library(tidyverse)
-# library(palmerpenguins)
 library(opendatatoronto)
@@
 These standards are applied to all apartment buildings with 10 or more units  or with 3 or more stories.
+(REF)
 
@@
 If a building gets a score of 50 or below, then then the full building will undergo a comprehensive inspection.
+(REF)
 
@@
 In a city like Toronto, where a lot of people do not own their personal property and live in rented apartments, it is critical that the landlord is kept informed of their building's condition so that they can ensure a tenants safety.
-Dhekte hein ki kuch badiya sa milta hai ya nahi inn socres ke baare mein.
-Aayie shuru karte, bina kisi rukawat ke.
 
-I think we should see which factors have the most effect on the evaluation score.
+What has been the results of introducing such a system? Has it been beneficial? Are there any specific features in a building that can give it more points and result in a better score?
 
@@
 
+We have taken Apartment Evaluation Score Data from Open Data Toronto(REF) and we will using this to gain a better understading of whether there have been anay improvements in the apartment building conditons.
+The data almost 10,000 entries spread over 40 columns.
+These columns give us a variety of information about each evaluation conducted, ranging from year built, year evaluated, property type, score, building address, wardnames, confirmed storeys, confimred units,stairwells, security rooms, laundry rooms, number of entrances, exits and a lot more.
+(REF)
+
+To analyze our dataset, I will be using R (REF), tidyverse (REF), dyplr (REF).
+There were multiple steps I took to clean and extract the neccesary data.
+The first thing I did was change the neccesssary column types from character to numeric, because initally columns such as SCORE and WARD were character type and it would not have been possible to graph them.
+Since we have a lot more entries of PRIVATE property type than PUBLIC and TCHC property types, I filtered out all entries that were not PRIVATE.
+Thus we will only be assesing the building evaluations for private apartment buildings.
+
+First I decide to take a look at the average scores of all the buildings based on the categories defined by the RentSafeTO Program (Ref).
+This meant categorizng scores according to the ranges they mentioned (Figure 1).
+To cateogirze the scores I added another column to the database that gave the cateogry based on their scores.
+I also reported the max, min, mean and standard deviation of all the scores in our dataset.
+(Table 1).
+Along with these two plots, I also made a bargraph based on the avereage scores of each of the 25 wards the apartment buildings are classified into.
+(Figure 2)
+
+The next four graphs are to see what helps buildings get a higher score than the rest.
+Since we have a lot of columns, I decided to focus on the most basic ones.
+Namely, Stairwells, Security, Laundry Rooms and Elevators.
+I made scatterplots using ggplot(REF) to see if any of these factors promote scores.
 ```{r}
@@
 data[,14:35] <- sapply(data[,14:35], as.numeric)
+data <- data %>% 
+  filter(PROPERTY_TYPE == "PRIVATE")
 
+mean_data <- data %>%
+  group_by(WARDNAME) %>%
+  mutate(mean_scores = mean(SCORE))
+
+a <- unique(mean_data$WARDNAME)
+b <- unique(mean_data$mean_scores)
+mean_wards <- data.frame(a, b)
 view(data)
+head(mean_wards)
+view(mean_wards)
 ```
 
-Talk more about it.
+```{r}
+compare_scores <- data %>%
+  as_tibble() %>%
+  select(WARD, SCORE) %>%
+  summarize(
+    The_min = min(SCORE),
+    The_max = max(SCORE),
+    Mean = mean(SCORE),
+    Std_dev = sd(SCORE)) %>%
+  arrange(desc(Mean))
 
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
+compare_scores %>%
+  knitr::kable(caption = "Variation of Ward Scores",
+               col.names = c("Score Min", "Score Max", "Score Mean", "Score Standard Deviation"),
+               align = c('l', 'l', 'l', 'l'),
+               booktabs = T)
+
+```
 
@@
 view(mut_data)
+
+#Figure 1
 mut_data %>%
-  ggplot(mapping = aes(x = score_category, fill = PROPERTY_TYPE)) +
+  ggplot(mapping = aes(x = score_category)) +
   geom_bar(position = "dodge")
@@
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+```{r}
+#Figure 2
+mean_wards %>%
+  ggplot(mapping = aes(x = b, y = a)) +
+  geom_bar(stat = "identity", position = "dodge")
 
-Equation \@ref(eq:bayes) seems useful, eh?
+```
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+```{r}
+#change fill
+#Figure 3
+data %>%
+  ggplot(mapping = aes(x = SCORE, y = STAIRWELLS, color = WARDNAME)) +
+  geom_point()
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+data %>%
+  ggplot(mapping = aes(x = SCORE, y = ELEVATORS, color = WARDNAME)) +
+  geom_point()
+
+data %>%
+  ggplot(mapping = aes(x = SCORE, y = SECURITY, color = WARDNAME)) +
+  geom_point()
+
+data %>%
+  ggplot(mapping = aes(x = SCORE, y = LAUNDRY_ROOMS, color = WARDNAME)) +
+  geom_point()
+
+```
 
@@
 
-## First discussion point
+## Initial Graphs
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The RentSafeTO bylaws say that a buildings next evaluation is determined based on their score.
+If the score is less than 50, then a complete buliding inspection and audit is conducted.
+This means that the building condition is not fit for living.
+The main aim of the program should be to see make sure each buiding provides a liveable space that is both safe for current residents and future ones.
+Figure 1 tells us that most of the residences fall in the 66-85 range, which means that their nexct evaluation comes 2 years later, This should be considered a good things as there a re very few apartments with scores lower than 50.
 
-## Second discussion point
+The table tells us that while some building did get a perfect score, some building also got zero points.
+While this is good to know, it is not helpful for our analysis.
+We do get to know that the average score is 72.4 which holds in line with our conclusion from the bargraph of categories.
 
-## Third discussion point
+Figure 2 shows us that while all wards are relatively close in scores, there are a few that on average better buildings than others.
+Don Valley North, Scarborough North, Scarborough Agincourt are the top three wards in terms of apartment quality.
+Davenport, Humber River-Black Creek and Parkdale- High Park are the wards with the three lowest wards.
+This gives us some insight as to how the residences are being maintained around Toronto, and can even help someone narrow down their search if they are looking for a place.
+
+## Scatterplots
+
+Our dataset consisted on multiple columns.
+Most of them told us the number of facilities in each apartment buliding.
+One would assume these features help a bulidings score, and these scatterplots give us some insight as to if this is true or not.
+Since there are a lot of facilites to account for, I decided to focus on the ones I find most important according to me, These plots do show an indication that having more is better.
+An increase in the number of stairwells, security, laundry rooms and elevators tend to give a higher evaluation scor, as indicated by Figure 3.
+
+While their are other factors included in the dataset, such as number of lobbies and garbage areas, these factors do not tend to play a crtical role in evaluations and these are also not things most people consider while looking at prospective residences.
+
+## Growth
+
+An important aspect of introducing and keeping large amounts on data on a program such as RentSafeTO in a populous and fast growing city like Toronto is to observe growth and see if it has been helpful or not.
+I calculated the averege scores over all the buildings and grouped them by the year of evaluation.
+The first year the prgram was conducted (2017), the average score was only 65.
+But since these were strict laws and building owners followed them, we saw a large incerease in the average score over the next 2 years, going up all the way to 79 in 2019.
+A few points have dropped in the last 2 years, but I think this can be attributed to the COVID-19 pandemic.
+
+```{r}
+#Table 2 
+year_mean_scores <- mut_data %>% 
+  group_by(YEAR_EVALUATED) %>%
+  summarise(value = mean(SCORE))
+year_mean_scores
+```
 
@@
 
-Weaknesses and next steps should also be included.
+A weakness that is evident can be seen in Table 2 above, which is the missing values.
+I could have removed the row from the table but it should be noted that this dataset contained a lot of missing cells in multiple rows and columns.
+A lot of columns are also not that helpful when you look at the way they are presented.
+The OTHER_FACILITIES column gives us the number of extra facilities certain luxury apartments might have.
+howver, since there is no description is provided, we have no idea of knowning what these features are.
+
+I belive a good step to make as this program continues is to maybe modify the cateogry values that are currently set.
+As we observed, the average score was 72 and the baseline is at 50.
+If this baseline is increased, the average score will go up and the quality of buildings all arund the city will increase.
+Adding some comments for each inspection will also help keep a track of any persisting problems that the buliding has not fixed over many evaluations.
 

### 2022-02-07 outputs/paper/paper.Rmd
added everything

@@
 ---
-title: "Toronto Apartment Evaluations Are Hard to Predict"
+title: "Apartment Evaluations Have Been Beneficial to the City of Toronto"
 #subtitle: "Subtitle Here"
@@
   bookdown::pdf_document2
-toc: FALSE
+toc: no
 bibliography: references.bib
@@
 library(opendatatoronto)
+library(scales)
+library(viridis)
+library(patchwork)
 ```
 
+\newpage
 # Introduction
 
-<!-- You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
--->
-
 The city of Toronto introcued a new bylaw enformenent program in July 2017 that is used to ensure that the apartment building owners and operators comply with the neccesary building maintanence standards, known as RentSafeTO.
-These standards are applied to all apartment buildings with 10 or more units  or with 3 or more stories.
-(REF)
+These standards are applied to all apartment buildings with 10 or more units  or with 3 or more stories.[@evaluations]
 
@@
 If the landlord does not comply to these requests, then according to the bylaws of RentSafeTO program, legal action can be taken against them.
+[@tenants]
 
@@
 If a building gets a score of 50 or below, then then the full building will undergo a comprehensive inspection.
-(REF)
+[@evaluations]
 
@@
 
-<!-- Our data is of penguins (Figure \@ref(fig:bills)). -->
-
-We have taken Apartment Evaluation Score Data from Open Data Toronto(REF) and we will using this to gain a better understading of whether there have been anay improvements in the apartment building conditons.
+We have taken Apartment Evaluation Score Data from Open Data Toronto [@toronto] and we will using this to gain a better understading of whether there have been anay improvements in the apartment building conditons.
 The data almost 10,000 entries spread over 40 columns.
 These columns give us a variety of information about each evaluation conducted, ranging from year built, year evaluated, property type, score, building address, wardnames, confirmed storeys, confimred units,stairwells, security rooms, laundry rooms, number of entrances, exits and a lot more.
-(REF)
+[@toronto]
 
-To analyze our dataset, I will be using R (REF), tidyverse (REF), dyplr (REF).
+To analyze our dataset, I will be using `R` [@citeR], `tidyverse` [@tidyverse], `dplyr` [@dplyr] along with `ggplot2` [@ggplot2], `scales` [@scales], `patchwork` [@patchwork] and `viridis` [@viridis].
 There were multiple steps I took to clean and extract the neccesary data.
@@
 
-First I decide to take a look at the average scores of all the buildings based on the categories defined by the RentSafeTO Program (Ref).
-This meant categorizng scores according to the ranges they mentioned (Figure 1).
+First I decide to take a look at the average scores of all the buildings based on the categories defined by the RentSafeTO Program [@evaluations].
+This meant categorizng scores according to the ranges they mentioned (Figure \@ref(fig:catscores)).
 To cateogirze the scores I added another column to the database that gave the cateogry based on their scores.
 I also reported the max, min, mean and standard deviation of all the scores in our dataset.
-(Table 1).
+(Table \@ref(fig:scoresummary)).
 Along with these two plots, I also made a bargraph based on the avereage scores of each of the 25 wards the apartment buildings are classified into.
@@
 Namely, Stairwells, Security, Laundry Rooms and Elevators.
-I made scatterplots using ggplot(REF) to see if any of these factors promote scores.
-```{r}
+I made scatterplots using `ggplot2`[@ggplot2] to see if any of these factors promote scores.
 
+```{r, include = FALSE}
 package <- show_package("4ef82789-e038-44ef-a478-a8f3590c3eb1")
@@
 mean_wards <- data.frame(a, b)
-view(data)
-head(mean_wards)
-view(mean_wards)
-```
-
-```{r}
-compare_scores <- data %>%
-  as_tibble() %>%
-  select(WARD, SCORE) %>%
-  summarize(
-    The_min = min(SCORE),
-    The_max = max(SCORE),
-    Mean = mean(SCORE),
-    Std_dev = sd(SCORE)) %>%
-  arrange(desc(Mean))
-
-compare_scores %>%
-  knitr::kable(caption = "Variation of Ward Scores",
-               col.names = c("Score Min", "Score Max", "Score Mean", "Score Standard Deviation"),
-               align = c('l', 'l', 'l', 'l'),
-               booktabs = T)
-
-```
-
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
-
-#unique proerty types
-unique_properties <- unique(data$PROPERTY_TYPE)
-
-data %>%
-  ggplot(mapping = aes(x = SCORE, y = CONFIRMED_UNITS, color= WARDNAME)) + 
-  geom_point(alpha = 0.5)
-
-data %>%
-  ggplot(mapping = aes(y = SCORE, fill=WARDNAME)) +
-  geom_bar()
 
@@
          )
-view(mut_data)
 
+view(data)
+head(mean_wards)
+view(mean_wards)
+```
+
+```{r, include=FALSE}
+compare_scores <- data %>%
+  as_tibble() %>%
+  select(WARD, SCORE) %>%
+  summarize(
+    The_min = min(SCORE),
+    The_max = max(SCORE),
+    Mean = mean(SCORE),
+    Std_dev = sd(SCORE)) %>%
+  arrange(desc(Mean))
+
+year_mean_scores <- mut_data %>% 
+  group_by(YEAR_EVALUATED) %>%
+  summarise(value = mean(SCORE))
+```
+
+```{r scoresummary,fig.cap="Variation of Apartment Evaluation Scores", echo=FALSE, warning=FALSE, message=FALSE}
+compare_scores %>%
+  knitr::kable(caption = "Variation of Ward Scores",
+               col.names = c("Score Min", "Score Max", "Score Mean", "Score Standard Deviation"),
+               align = c('l', 'l', 'l', 'l'),
+               booktabs = T)
+```
+
+<!-- \vspace{12pt} -->
+```{r catscores, fig.cap="Number of buildings in each of the 4 RentSafeTO evaluation categories",echo = FALSE}
 #Figure 1
-mut_data %>%
-  ggplot(mapping = aes(x = score_category)) +
-  geom_bar(position = "dodge")
+ggplot(mut_data,mapping = aes(x = score_category)) +
+geom_bar(position = "dodge", fill = "steelblue") + 
+theme_minimal() +
+labs(x = "RentSafeTO categories for apartment scores",
+     y = "Number of buildings")
 ```
@@
 
-# Model
+# Models
 
-```{r}
+```{r wardmeans, fig.cap= "Variation of the average scores for each ward", echo=FALSE, warning=FALSE, message=FALSE}
 #Figure 2
 mean_wards %>%
-  ggplot(mapping = aes(x = b, y = a)) +
-  geom_bar(stat = "identity", position = "dodge")
+  ggplot(mapping = aes(x = a, y = b, fill= a)) +
+  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
+  labs(y = "Average Evaluation Scores",
+       x = "Ward Names") +
+  theme(axis.text.x = element_text(angle = 65, hjust = 1),
+        legend.position = "none") +
+#  scale_fill_gradient(low = "white", high = "black") +
+  scale_y_continuous(limits=c(65,80), oob = rescale_none)
 
@@
 
-```{r}
+```{r stairs, fig.cap="Relationship between number of stairwells and score",echo=FALSE, warning=FALSE, message=FALSE, out.width='80%'}
 #change fill
@@
 data %>%
-  ggplot(mapping = aes(x = SCORE, y = STAIRWELLS, color = WARDNAME)) +
-  geom_point()
-
+  ggplot(mapping = aes(x = SCORE, y = STAIRWELLS, color = STAIRWELLS)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Evaluation Score",
+       y = "Number of Stairwells",
+       caption = "Relationship between increase in Stairwells and Evaluation Score")+ 
+  scale_color_viridis(discrete = FALSE, option = "D")+
+  scale_fill_viridis(discrete = FALSE) 
+```
+```{r elevators, fig.cap="Relationship between number of elevators and score",echo=FALSE, warning=FALSE, message=FALSE, out.width='80%'}
+#change fill
+#Figure 3
 data %>%
-  ggplot(mapping = aes(x = SCORE, y = ELEVATORS, color = WARDNAME)) +
-  geom_point()
-
+  ggplot(mapping = aes(x = SCORE, y = ELEVATORS, color = ELEVATORS)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Evaluation Score",
+       y = "Number of Elevators",
+       caption = "Relationship between increase in Elevators and Evaluation Score")+ 
+  scale_color_viridis(discrete = FALSE, option = "D")+
+  scale_fill_viridis(discrete = FALSE) 
+```
+```{r security, fig.cap="Relationship between number of security and score",echo=FALSE, warning=FALSE, message=FALSE, out.width='80%'}
+#change fill
+#Figure 3
 data %>%
-  ggplot(mapping = aes(x = SCORE, y = SECURITY, color = WARDNAME)) +
-  geom_point()
-
+  ggplot(mapping = aes(x = SCORE, y = SECURITY, color = SECURITY)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Evaluation Score",
+       y = "Number of Security",
+       caption = "Relationship between increase in Security and Evaluation Score")+ 
+  scale_color_viridis(discrete = FALSE, option = "D")+
+  scale_fill_viridis(discrete = FALSE) 
+```
+```{r laundry, fig.cap="Relationship between number of laundry rooms and score",echo=FALSE, warning=FALSE, message=FALSE, out.width='80%'}
+#change fill
+#Figure 3
 data %>%
-  ggplot(mapping = aes(x = SCORE, y = LAUNDRY_ROOMS, color = WARDNAME)) +
-  geom_point()
-
+  ggplot(mapping = aes(x = SCORE, y = LAUNDRY_ROOMS, color = LAUNDRY_ROOMS)) +
+  geom_point() +
+  theme_minimal() +
+  labs(x = "Evaluation Score",
+       y = "Number of Laundry Rooms",
+       caption = "Relationship between increase in Laundry Rooms and Evaluation Score")+ 
+  scale_color_viridis(discrete = FALSE, option = "D")+
+  scale_fill_viridis(discrete = FALSE) 
 ```
 
-# Results
+\newpage
 
@@
 
-```{r}
+```{r yearavg,fig.cap="Variation of Apartment Evaluation Year-wise", echo=FALSE, warning=FALSE, message=FALSE}
 #Table 2
-year_mean_scores <- mut_data %>% 
-  group_by(YEAR_EVALUATED) %>%
-  summarise(value = mean(SCORE))
-year_mean_scores
+
+year_mean_scores %>%
+  knitr::kable(caption = "Yearly Score Changes",
+               col.names = c("Year", "Average Score"),
+               align = c('l','l'),
+               booktabs = T)
 ```
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional details
-
-\newpage
-
 # References
-https://www.toronto.ca/community-people/housing-shelter/rental-housing-tenant-information/rental-housing-standards/apartment-building-standards/rentsafeto-for-tenants/
-
-https://www.toronto.ca/community-people/housing-shelter/rental-housing-tenant-information/rental-housing-standards/apartment-building-standards/rentsafeto-for-building-owners/rentsafeto-building-evaluations-and-audits/

### 2022-02-07 outputs/paper/paper.Rmd
finishing touches

@@
          )
-
-view(data)
-head(mean_wards)
-view(mean_wards)
 ```

### 2022-02-07 outputs/paper/paper.Rmd
rename

@@
   # - Another author
-thanks: "Code and data are available at: www.github.com/ritvikpuri/sta304-paper-1"
+thanks: "Code and data are available at: www.github.com/ritvikpuri/Toronto-Apartment-Evaluations"
 date: "`r format(Sys.time(), '%d %B %Y')`"

### 2022-02-07 outputs/paper/paper.Rmd
accesory work

@@
 
+#changing data types to numeric
 sapply(data, class)
@@
 
+#adding a mean data column for each ward
 mean_data <- data %>%
@@
 
+# Conclusion
+
+Whenever the government introduces a new program or policy, I belive it has higher chances of failing than success.
+Especially if its something new that people don't expect and take time getting accustomed to.
+The RentSafeTO program according to me turned out to be a successfull one.
+We can clearly see a growth in the average score every year, albeit taking a bit of a hit due the global coronavirus pandemic in the last 2 years.
+This growth means that the living conditions have gone from good to better, and more people are inclined to get a place in the city.
+After things normalize from the pandemic, continuing this program will make sure building owners and landlords abide to it and make sure to maintain a safe and healthy place for poeple to live in.
+I hope the living conditions continue to get better through the help of this program, even if its just a small contribution.
+
 \newpage

### 2022-02-07 scripts/01-data_cleaning.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Clean the survey data downloaded from https://open.toronto.ca/dataset/apartment-building-evaluation/
+# Author: Ritvik Puri
+# Data: 6 February 2022
+# Contact: ritvik.puri@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
 # - Need to have downloaded the ACS data and saved it to inputs/data
@@ -27,9 +27,6 @@ reduced_data <-
   select(first_col, 
          second_col)
 rm(raw_data)
-         
-
-#### What's next? ####
 
 
 
